Combined Task and Motion Planner through an Interface Layer

### 2015-05-09 documentation/ps4.tex

The interface layer starts by asking the task planner to generate a high level plan and has the motion planner instantiate motion plans corresponding to each action.  
When it's not possible to generate a motion plan, the interface layer determines what geometric constraints are causing the problem and relays this back to the task planner.  
Then, the task planner replans from that state and the process continues.  
When necessary, backtracking occurs.

At start time, the task planner is given a PDDL domain specifying the different actions that are feasible along with their preconditions and postconditions.  
For individual calls to the task planner, the interface layer takes an internal `state' variable, converts it to PDDL format, and sends it along to be solved.  
The task planner responds with a high level plan or a notification that finding a plan failed.  
If finding a plan failed, the interface layer begins backtracking.  
If a high level plan is found, the interface layer begins searching for refinements of each action as a motion plan.

@@

\item Call the task planner to create a high level plan.  
In this case, the task planner doesn't yet know that block 2 is blocking block 1, so the plan is simply:

@@

\item Next step is to call the motion planner to try and find a trajectory corresponding to each of these actions.  
In this case, the motion planner fails to find a motion plan for step 1.

\item The interface layer goes through removing objects to try to find what was causing the error for that action.  
The interface layer discovers that if it removes block 2, then the motion plan can be created just fine, so block 2 must be obstructing the path.

@@

\item Now, we try to find a new high level plan.  
Since block 2 is obstructing block 1, we can't pick up block 1 directly, so the task planner realizes it needs to move block 2 first before moving block 1. 
The new plan it creates is:

@@

One of the key insights of this paper is that continuous variables can be referenced symbolically in the task planner within the PDDL domain.  
For instance, consider the pose of the robot: this is a continuous, geometric state.  However, the task planner only keeps track of poses symbolically.  
The task planner knows that there is some GPFG\_obj\_1 -- a Gripper Pose For Grasping -- for object 1, and it doesn't really care what that pose is, so long as it exists.  
The task planner reasons over a variety of these symbolic variables, corresponding to poses and trajectories.  
The motion planner and pose generators take these symbolic variables and associate them with corresponding real locations and motions in the world.  
This is the key insight which allows the task planner to reason about continuous variables without discretizing.

We will follow [1] in using manipulation as an example environment.  For any planning domain with some continuous variables and some discrete variables, it will be possible to transform the continuous variables to discrete symbolic references to continuous positions.  For this example, we talk through the PICKUP action for a simple manipulator; the PUTDOWN action is analagous.

@@

Here, we have three continuous variables that we represent symbolically: $traj$, for each of the trajectories, and $pose_1$ and $pose_2$, representing start and end locations for the robot's gripper.
The preconditions and effects can also iterate over all of these symbolic variables, allowing us to mark that a picked up object doesn't obstruct any other objects, etc. 
We have various predicates which describe different things about both the discrete variables and the continuous variables.

@@

How do we come up with all of these symbolic references?  
We create as many as we need.  
In this case, we need to have a pose where we can pick up each object and put each object down.  
Neither the interface layer nor the task planner know what the poses correspond to in the real world.
The pose generators instantiate specific poses which the motion planner verifies or discounts if it doesn't work.
Similarly, we have one trajectory for each potential pair of poses.  
However, since the trajectory isn't actually used in any of the effects, we actually can ignore it entirely -- it doesn't change the high level plan and the interface layer will still know to have the motion planner create the relevant trajectories. 
So to make the plan more efficient, we actually cut out the $ISMP$ predicate and all references to trajectories.

@@

When the pose generator is called by the interface layer, it generates a set of random poses for each action. Each pose is made up of a set of waypoints. For example for the PICKUP action, it generates the following waypoints: near the cylinder and pointing towards it, touching the cylinder, lifting the cylinder, and a standard home position. It passes one pose to the interface layer at a time.  if none of the poses are feasible, the interface layer backtracks and resets the pose generator. The TryRefine method of the interface layer attempts to generate a feasible motion plan between each pose.  The motion planner is called to create this motion plan, including motion plans between waypoints.

@@

This figure (Fig. \ref{fig:figure4}), included with text from [1], explains the flow of control of the interface layer through a variety of situations, including more than the simple case covers.  
The subfigures in this diagram explain on a high level the different steps the interface layer takes when solving a problem.  
The subfigure a shows the motion planner's backtracking search when trying to instantiate motions for each of the high level actions.  
Subfigure b shows the first possible scenario, where the backtracking search succeeds.  Subfigure c shows a second possible scenario, where the backtracking search doesn't succeed.  
There are multiple options of what can happen within this scenario: subfigure d shows an option where they determine that the block2 obstructs the trajectory.  
This triggers the behavior discussed earlier, where the state is updated for the task planner.  
Subfigure e shows the new task plan after replanning occurs.  
After this, the interface layer will repeat the process of attempting to refine this new plan.  
Subfigure f shows an alternative option, where the blocking object is determined to be the table, which is not movable.  
Then, the state is updated as before and the task planner is called, but no solution is found, so backtracking occurs.

@@

The original version of the algorithm, as shown in (Fig. \ref{fig:algorithms}), only attempts to find partial trajectories when it replans partway through the procedure.  Our algorithm has a slight modification. Every time a new partial plan is generated, we first try to fill it completely using backtracking search before giving up and resorting to the partial trajectory finding.  In some scenarios, some poses will not work but others will, so it is better to try to make the simplest possible plan work before giving up and making the high level plan more complicated.  In our manipulation example, this turns into trying to pick up blocks from all angles before giving up and assuming they must be blocked by something.  An analysis of the benefits of this approach are outlined in the benchmarking section.

@@

If that doesn't succeed, we find a partial trajectory that goes as far as it can by calling the TryRefine subroutine in partial trajectory mode.  
This function returns a partial trajectory along with the step where it failed and the determined cause for failure.  
Then, we update the state based on the actions up until that point and add in the obstructing obstacles corresponding to the failure.  
Now, we attempt to find a new high level plan, and iterate.

@@

To find the blocking objects, we start by iterating through a randomized list of the objects.  We continue removing objects until we find a successful motion plan.  
We will eventually find a successful motion plan since if there are no other objects in the world, the motion should succeed.  
After we find this possible list of removed objects, we iterate through all of the objects we removed.  
We attempt to add these objects in one by one, leaving added in all of those which don't obstruct our found motion plan.  
This leaves us with a local minimum of removed objects: it may be possible to find another smaller set of objects that, if removed, allow a valid motion plan.  
However, for this set of this set of objects, no object can be added back in without obstructing the motion plan.

@@

\section{Benchmarking}
\subsection{Testing}
\begin{figure}[t]
\centering
\def\svgwidth{0.5\textwidth}
\input{figures/smallEnv.pdf_tex}
\caption{Small Environment Test Example. The goal was for the blue blocks to be on surface I and the red blocks to be on surface S.\label{fig:smallEnv}}
\end{figure}
We generated a set of small-scale tests to check the correctness of the interface layer. ref{fig-smallTest} Each test had three to four blocks initially placed on one or two surfaces. Some blocks were surrounded on 3 sides and others on all sides by walls and blocks. For blocks surrounded on 3 sides, the interface layer had to find the pose that picked up the block in the correct direction. For blocks surrounded on all sides, the interface layer needed to generate actions to move blocks in the way.

\subsection{Procedure}
\begin{figure}[t]
\centering
\def\svgwidth{\textwidth}
\input{figures/benchmarks.pdf_tex}
\caption{Benchmarking Test Examples. The goal was to move the green block to another surface.\label{fig:benchmark}}

@@

We performed a variety of benchmarking tests to measure the performance of the interface. We measured the performance of each test by the number of calls to both the task planner and motion planner and the runtime. ref{table-benchmarkTable}
The difficulty of a benchmark increased as the total number \textit{obstructions} increased. The total number of obstructions was counted by summing the number of directions around each block that were obstructed. For the ordinal directions, a neighboring block in any one of three locations obstructs the direction. For example, a neighboring block above, to the right or to the top right corner of a block obstructs the northeast direction of the block. Furthermore, as the number of blocks for a particular grid shape increased, the test difficulty increased, since the total number of obstructions increased.

\subsection{Results}
\begin{figure}[t]
	\centering
	\subfloat{\includegraphics[width=.45\linewidth]{figures/bvm}}
	\qquad
	\subfloat{\includegraphics[width=.45\linewidth]{figures/bvt}}
	\qquad
	\subfloat{\includegraphics[width=.45\linewidth]{figures/bvrt}}
	\qquad
	\subfloat{\includegraphics[width=.45\linewidth]{figures/ovm}}
	\qquad
	\subfloat{\includegraphics[width=.45\linewidth]{figures/ovt}}
	\qquad
	\subfloat{\includegraphics[width=.45\linewidth]{figures/ovrt}}
	\qquad
	\subfloat{\includegraphics[width=.45\linewidth]{figures/square}}
	\caption{Benchmarking plots\label{fig:benchmarkPlots}}
\end{figure}
The goal of each test was to move the green block, at the center of a grid of blocks, to a goal surface. 
We tested a variety of different grid shapes, including a square grid, an X, a cross, and lines. 
We tested each grid shape over a range of sizes ranging from 3 to 13. ref{fig-benchmarks} 
The most difficult benchmark we tested was the 13x13 square grid. Larger tests caused the FF planner to segfault. As seen in \ref{fig:benchmarkPlots}, as the size of the problem increases the number of calls increases.

Figure \ref{benchmark2} shows the results for the original algorithm, minus our modification.  (Our modification always tries to find a perfect plan after adding a new high level plan section to the plan, rather than immediately trying to find a new partial trajectory.)  For the very simplest cases, the original algorithm performs better, since it makes the same number of motion and task planner calls but has less backtracking overhead.  However, for the more complicated environments, the cross and the square, our algorithm was significantly faster, as shown in \ref{benchmark}.

The interface layer is designed to be independent of any specific task and motion planner libraries, therefore number of calls to the motion and task planner measures the performance of the interface layer better than the total runtime, since the actual runtime will be very dependent on the motion planner and task planner chosen.

\begin{figure}
\begin{tabular}[t]{|l|l|l|l|l|l|}
\hline
Shape & Blocks & obstructions & Motion Planner Calls & Task Planner Calls & runtime \\ \hline
Horizontal Line & 3 & 12 & 3 & 1 & .023\\
& 5 & 24 & 4 & 1 & .028\\
& 7 & 36 & 4 & 1 & .029\\
& 9 & 48 & 4 & 1 & .029\\
& 11 & 60 & 4 & 1 & .034\\
& 13 & 72 & 4 & 1 & .037\\
\hline
Vertical Line & 3 & 12 & 4 & 1 & .024\\
& 5 & 24 & 4 & 1 & .027\\
& 7 & 36 & 4 & 1 & .028\\
& 9 & 48 & 4 & 1 & .028\\
& 11 & 60 & 4 & 1 & .037\\
& 13 & 72 & 4 & 1 & .036\\ 
\hline
X & 5 & 8 & 2 & 1 & .026\\
& 9 & 16 & 2 & 1 & .034\\
& 13 & 24 & 2 & 1 & .036\\
& 17 & 32 & 2 & 1 & .047\\
& 21 & 40 & 2 & 1 & .066\\
& 25 & 48 & 2 & 1 & .079\\ 
\hline
Cross & 5 & 20 & 39 & 3 & .16\\
& 9 & 44 & 50 & 2 & .20\\
& 13 & 68 & 59 & 2 & .23\\
& 17 & 92 & 71 & 3 & .26\\
& 21 & 116 & 78 & 3 & .31\\
& 25 & 140 & 77 & 3 & .39\\ 
\hline
Square & 9 & 56 & 54 & 2 & .16\\
& 25 & 176 & 294 & 6 & 1.06\\
& 49 & 360 & 818 & 10 & 5.07\\
& 81 & 608 & 3177 & 24 & 34.68\\
& 121 & 920 & 3681 & 22 & 87.06\\
& 169 & 1296 & 14216 & 57 & 518.61\\ 
\hline

\end {tabular}
\caption{Benchmark Tests - Our Modified Algorithm}
\label{benchmark}
\end{figure}

\begin{figure}
\begin{tabular}[t]{|l|l|l|l|l|l|}
\hline
Shape & Blocks & obstructions & Motion Planner Calls & Task Planner Calls & runtime \\ \hline
Horizontal Line & 3 & 12 & 3 & 1 & .016\\
& 9 & 48 & 4 & 1 & .025\\
& 13 & 72 & 4 & 1 & .033\\
\hline
Vertical Line & 3 & 12 & 4 & 1 & .022\\
& 9 & 48 & 4 & 1 & .027\\
& 13 & 72 & 4 & 1 & .031\\ 
\hline
X & 5 & 8 & 2 & 1 & .014\\
& 17 & 32 & 2 & 1 & .031\\
& 25 & 48 & 2 & 1 & .056\\ 
\hline
Cross & 5 & 20 & 47 & 5 & .15\\
& 17 & 92 & 206 & 10 & .68\\
& 25 & 140 & 390 & 14 & 1.56\\ 
\hline
Square & 9 & 56 & 145 & 11 & .45\\
& 25 & 176 & 4354 & 143 & 16.55\\

\hline

\end {tabular}
\caption{Benchmark Tests - Original Algorithm}
\label{benchmark2}
\end{figure}
\section{Limitations}

@@

The two immediately relevant open questions are whether it's possible to design a similar interface between task and motion plans which returns optimal plans or which can be used to do online planning.
Currently, the interface layer works only in deterministic, fully observable environments.  
It would be very interesting to develop similar algorithms to work in non-deterministic or partially observable environments as well.

>>>>

The interface layer starts by asking the task planner to generate a high level plan and has the motion planner instantiate motion plans corresponding to each action.  
When it is not possible to generate a motion plan, the interface layer determines what geometric constraints are causing the problem and relays this back to the task planner.  
Then, the task planner replans from that state and the process continues.  
When necessary, backtracking occurs.

At start time, the task planner is given a PDDL domain specifying the different actions that are feasible along with their preconditions and postconditions.  
For individual calls to the task planner, the interface layer takes an internal \textit{state} variable, converts it to PDDL format, and sends it along to be solved.  
The task planner responds with a high level plan or a notification that finding a plan failed.  
If finding a plan failed, the interface layer begins backtracking.  
If a high level plan is found, the interface layer begins searching for refinements of each action as a motion plan.

@@

\item Call the task planner to create a high level plan.  
In this case, the task planner has not yet inferred that block 2 is blocking block 1, so the plan is simply:

@@

\item The next step is to call the motion planner to try to find a trajectory corresponding to each of these actions.  
In this case, the motion planner fails to find a motion plan for step 1.

\item The interface layer removes objects to try to figure out which one was causing the error for that action.  
The interface layer discovers that if it removes block 2, then the motion plan can be created just fine, so block 2 must be obstructing the path.

@@

\item Now, we try to find a new high level plan.  
Since block 2 is obstructing block 1, we cannot pick up block 1 directly, so the task planner realizes it needs to move block 2 first before moving block 1. 
The new plan it creates is:

@@

One of the key insights of this paper is that continuous variables can be referenced symbolically in the task planner within the PDDL domain.  
For instance, consider the pose of the robot: this is a continuous, geometric state.
However, the task planner only keeps track of poses symbolically.  
The task planner knows that there is some GPFG\_obj\_1 -- a Gripper Pose For Grasping -- for object 1, and it does not really care what that pose is, so long as it exists.  
The task planner reasons over a variety of these symbolic variables, corresponding to poses and trajectories.  
The motion planner and pose generators take these symbolic variables and associate them with corresponding real locations and motions in the world.  
This is the key insight which allows the task planner to reason about continuous variables without discretizing.

We follow [1] in using manipulation as an example environment.  For any planning domain with some continuous variables and some discrete variables, it is possible to transform the continuous variables to discrete symbolic references to continuous positions.  For this example, we talk through the PICKUP action for a simple manipulator; the PUTDOWN action is analagous.

@@

Here, we have three continuous variables that we represent symbolically: $traj$, for each of the trajectories, and $pose_1$ and $pose_2$, representing start and end locations for the robot's gripper.
The preconditions and effects can also iterate over all of these symbolic variables, allowing us to mark that a picked up object does not obstruct any other objects, etc. 
We have various predicates which describe different things about both the discrete variables and the continuous variables.

@@

How do we come up with all of these symbolic references?  
We create as many as we need.  
In this case, we need to have a pose where we can pick up each object and put each object down.  
Neither the interface layer nor the task planner know what the poses correspond to in the real world.
The pose generators instantiate specific poses which the motion planner verifies or discounts if it fails.
Similarly, we have one trajectory for each potential pair of poses.  
However, since the trajectory is not actually used in any of the effects, we actually can ignore it entirely -- it does not change the high level plan and the interface layer will still know to have the motion planner create the relevant trajectories. 
So to make the plan more efficient, we actually cut out the $ISMP$ predicate and all references to trajectories.

@@

When the pose generator is called by the interface layer, it generates a set of random poses for each action. 
Each pose is made up of a set of waypoints. For example for the PICKUP action, it generates the following waypoints: near the cylinder and pointing towards it, touching the cylinder, lifting the cylinder, and a standard home position. 
It passes one pose to the interface layer at a time.  
If none of the poses are feasible, the interface layer backtracks and resets the pose generator. 
The TryRefine method of the interface layer attempts to generate a feasible motion plan between each pose.  The motion planner is called to create this motion plan, including motion plans between waypoints.

@@

This figure (Fig. \ref{fig:figure4}), included with text from [1], explains the flow of control of the interface layer through a variety of situations, including more than the simple case covers.  
The subfigures in this diagram explain on a high level the different steps the interface layer takes when solving a problem.  
The subfigure a shows the motion planner's backtracking search when trying to instantiate motions for each of the high level actions.  
Subfigure b shows the first possible scenario, where the backtracking search succeeds.  Subfigure c shows a second possible scenario, where the backtracking search does not succeed.  
There are multiple options of what can happen within this scenario: subfigure d shows an option where they determine that the block2 obstructs the trajectory.  
This triggers the behavior discussed earlier, where the state is updated for the task planner.  
Subfigure e shows the new task plan after replanning occurs.  
After this, the interface layer will repeat the process of attempting to refine this new plan.  
Subfigure f shows an alternative option, where the blocking object is determined to be the table, which is not movable.  
Then, the state is updated as before and the task planner is called, but no solution is found, so backtracking occurs.

@@

The original version of the algorithm, as shown in (Fig. \ref{fig:algorithms}), only attempts to find partial trajectories when it replans partway through the procedure.
Our algorithm has a slight modification. Every time a new partial plan is generated, we first try to fill it completely using backtracking search before giving up and resorting to the partial trajectory finding.  
In some scenarios, some poses will not work but others will, so it is better to try to make the simplest possible plan work before giving up and making the high level plan more complicated.  
In our manipulation example, this turns into trying to pick up blocks from all angles before giving up and assuming they must be blocked by something.  
An analysis of the benefits of this approach are outlined in the results section.

@@

If that does not succeed, we find a partial trajectory that goes as far as it can by calling the TryRefine subroutine in partial trajectory mode.  
This function returns a partial trajectory along with the step where it failed and the determined cause for failure.  
Then, we update the state based on the actions up until that point and add in the obstructing obstacles corresponding to the failure.  
Now, we attempt to find a new high level plan, and iterate.

@@

To find the blocking objects, we start by iterating through a randomized list of the objects.  We continue removing objects until we find a successful motion plan.  
We will eventually find a successful motion plan since if there are no other objects in the world, the motion should succeed.  
After we find this possible list of removed objects, we iterate through all of the objects we removed.  
We attempt to add these objects in one by one, leaving added in all of those which do not obstruct our found motion plan.  
This leaves us with a local minimum of removed objects: it may be possible to find another smaller set of objects that, if removed, allow a valid motion plan.  
However, for this set of this set of objects, no object can be added back in without obstructing the motion plan.

@@

\section{Results}
\subsection{Test Scenarios}
\begin{figure}[t]
\centering
\def\svgwidth{\textwidth}
\input{figures/scenarios.pdf_tex}
\caption{Test Examples. The goal is to move the green block to another surface.\label{fig:scenarios}}

@@

Baxter executes a variety of tests to measure the performance of the interface. 
The goal of each test is to move the cylinder at the center of an arrangment of objects, to a goal surface. 
We tested a variety of different 3x3 object arrangements on Baxter, including a square grid, an X, a cross, a vertical line, and a horizontal line [Fig. \ref{fig:scenarios}].

The most difficult scenario we tested was the square grid.
The difficulty of a scenario depends on the number of blocks for a particular grid shape. 
As the total number of \textit{obstructions} increases, the test difficulty increased. 
The total number of obstructions is measured by counting the number of directions around each object that were obstructed. 
For the ordinal directions, a neighboring object in any one of three locations obstructs the direction. 
For example, the northeast direction of an object is obstructed by a neighboring object above, to the right or to the top right corner.

During preliminary testing, the results for the original algorithm took significantly longer when tested in simulation.  
Our modification always tries to find a perfect plan after adding a new high level plan section to the plan, rather than immediately trying to find a new partial trajectory. 
Since our algorithm was significantly faster we used it for each of the test scenarios.

We measured the performance of each test by the runtime and the number of calls to both the task planner and motion planner [Fig. \ref{benchmark}]. 
The interface layer is designed to be independent of any specific task and motion planner libraries, therefore number of calls to the motion and task planner measures the performance of the interface layer better than the total runtime, since the actual runtime will be very dependent on the motion planner and task planner chosen.

\subsection{Test Results}

\begin{figure}
\begin{tabular}[t]{|l|l|l|l|l|l|}
\hline
Shape & Blocks & obstructions & Motion Planner Calls & Task Planner Calls & runtime \\ \hline
Horizontal Line & 3 & 12 & ? & ? & ?\\
\hline
Vertical Line & 3 & 12 & ? & ? & ?\\
\hline
X & 5 & 8 & ? & ? & ?\\
\hline
Cross & 5 & 20 & ? & ? & ?\\
\hline
Square & 9 & 56 & ? & ? & ?\\
\hline

\end {tabular}
\caption{Test Scenarios Performance}
\label{benchmark}
\end{figure}

\section{Limitations}

@@

The two immediately relevant open questions are whether it is possible to design a similar interface between task and motion plans which returns optimal plans or which can be used to do online planning.
Currently, the interface layer works only in deterministic, fully observable environments.  
It would be very interesting to develop similar algorithms to work in non-deterministic or partially observable environments as well.
